A highly-available move operation for replicated trees

### 2019-06-12 paper/move-op.tex

\usepackage{lineno}
\usepackage{isabelle,isabellesym}
\isabellestyle{tt}
\renewenvironment{isabelle}{%
  \bigbreak\noindent\hspace{\parindent}%
  \begin{minipage}{\textwidth-\parindent}% use minipage to prevent page breaks
  \begin{isabellebody}%
  \begin{tabbing}%
}{%
  \end{tabbing}%
  \end{isabellebody}%
  \end{minipage}%
  \bigbreak%
}
\renewcommand{\isacartoucheopen}{}
\renewcommand{\isacartoucheclose}{}

@@

\ \ {\isacartoucheopen}unique{\isacharunderscore}parent\ tree\ {\isasymequiv}\ {\isacharparenleft}{\isasymforall}p{\isadigit{1}}\ p{\isadigit{2}}\ m{\isadigit{1}}\ m{\isadigit{2}}\ c{\isachardot}\ {\isacharparenleft}p{\isadigit{1}}{\isacharcomma}\ m{\isadigit{1}}{\isacharcomma}\ c{\isacharparenright}\ {\isasymin}\ tree\ {\isasymand}\ {\isacharparenleft}p{\isadigit{2}}{\isacharcomma}\ m{\isadigit{2}}{\isacharcomma}\ c{\isacharparenright}\ {\isasymin}\ tree\ {\isasymlongrightarrow}\ p{\isadigit{1}}\ {\isacharequal}\ p{\isadigit{2}}\ {\isasymand}\ m{\isadigit{1}}\ {\isacharequal}\ m{\isadigit{2}}{\isacharparenright}{\isacartoucheclose}%

@@

>>>>

\usepackage{lineno}
\usepackage{isabelle,isabellesym}
\isabellestyle{tt}
\renewenvironment{isabelle}{%
  \bigbreak\noindent%
  \renewcommand{\isanewline}{\\}%
  \begin{minipage}{\columnwidth}% use minipage to prevent page breaks
  \begin{isabellebody}%
  \begin{tabbing}%
}{%
  \end{tabbing}%
  \end{isabellebody}%
  \end{minipage}%
  \bigbreak%
}
\renewcommand{\isacartoucheopen}{}
\renewcommand{\isacartoucheclose}{}

@@

\ \ {\isacartoucheopen}unique{\isacharunderscore}parent\ tree\ {\isasymequiv}\ {\isacharparenleft}{\isasymforall}p{\isadigit{1}}\ p{\isadigit{2}}\ m{\isadigit{1}}\ m{\isadigit{2}}\ c{\isachardot}\ {\isacharparenleft}p{\isadigit{1}}{\isacharcomma}\ m{\isadigit{1}}{\isacharcomma}\ c{\isacharparenright}\ {\isasymin}\ tree\ {\isasymand}\ {\isacharparenleft}p{\isadigit{2}}{\isacharcomma}\ m{\isadigit{2}}{\isacharcomma}\ c{\isacharparenright}\ {\isasymin}\ tree\ {\isasymlongrightarrow}\ p{\isadigit{1}}\ {\isacharequal}\ p{\isadigit{2}}\ {\isasymand}\ m{\isadigit{1}}\ {\isacharequal}\ m{\isadigit{2}}{\isacharparenright}{\isacartoucheclose}\isanewline%
\ acyclic\ {\isacharcolon}{\isacharcolon}\ {\isacartoucheopen}{\isacharparenleft}{\isacharprime}n\ {\isasymtimes}\ {\isacharprime}m\ {\isasymtimes}\ {\isacharprime}n{\isacharparenright}\ set\ {\isasymRightarrow}\ bool{\isacartoucheclose}\ \isakeyword{where}\isanewline
\ \ {\isacartoucheopen}acyclic\ tree\ {\isasymequiv}\ {\isacharparenleft}{\isasymnexists}n{\isachardot}\ ancestor\ tree\ n\ n{\isacharparenright}{\isacartoucheclose}

@@

We now discuss the correctness properties of the algorithm from Section~\ref{sec:algorithm}.
All theorems stated here have been formally proved and mechanically checked using Isabelle.
For space reasons we give only a brief sketch of the proofs in this paper, but the full details can be found in the Isabelle files of our open source proof development.\footnote{\url{https://github.com/trvedata/move-op}}

\subsection{Tree invariants}\label{sec:tree-invariants}

A tree is an acyclic graph in which every node has exactly one parent, except for the root, which has no parent.
To prove that our algorithm always maintains a tree structure, no matter which operations are applied, we need to prove that these invariants hold.

In fact, we slightly generalise this property: we allow more than one root to exist, so the graph is actually a forest.
This allows an application to move nodes between multiple trees, if desired.
For example, the ``trash'' node used for deletion (Section~\ref{sec:ops-trees}) can be separate from the main tree.
If the

\begin{isabelle}
\isacommand{theorem}\isamarkupfalse%
\ apply{\isacharunderscore}ops{\isacharunderscore}unique{\isacharunderscore}parent{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}apply{\isacharunderscore}ops\ ops\ {\isacharequal}\ {\isacharparenleft}log{\isacharcomma}\ tree{\isacharparenright}{\isacartoucheclose}\isanewline
\ \ \isakeyword{shows}\ {\isacartoucheopen}unique{\isacharunderscore}parent\ tree{\isacartoucheclose}
\end{isabelle}

\begin{isabelle}
\isacommand{theorem}\isamarkupfalse%
\ apply{\isacharunderscore}ops{\isacharunderscore}acyclic{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}apply{\isacharunderscore}ops\ ops\ {\isacharequal}\ {\isacharparenleft}log{\isacharcomma}\ tree{\isacharparenright}{\isacartoucheclose}\isanewline
\ \ \isakeyword{shows}\ {\isacartoucheopen}acyclic\ tree{\isacartoucheclose}
\end{isabelle}
